\documentclass{paper.tex}[subfiles]
\begin{document}

\section{Background}

Most continuous mathematics uses the real number system,
  but this system cannot be represented faithfully
  in a bounded number of bits.
Computer programs that aim to represent continuous mathematics
  instead use floating point arithmetic.
Since this representation cannot perfectly match real numbers,
  programmers who use floating point numbers
  must avoid a variety of subtle misbehaviors
  lest the program compute answers
  unrelated to the real-number computation it attempts to model.
This section describes floating point numbers
  and the difficulties in using them to describe
  real-number continuous calculations.

\subsection{Floating point numbers}

Floating point numbers represent a subset of real numbers,
  distributed approximately exponentially, the numbers of the form
\[ \pm (1 + m) 2^e, \]
  where $m$, the \emph{mantissa}, is a fixed-width number in $[0, 1]$,
  and $e$, the \emph{exponent}, is a fixed-width signed integer.
Normal floating point numbers are thus represented by
  a sign bit, exponent, and mantissa.
Floating point can also represent a few special values:
  \emph{denormal} numbers of form $\pm m 2^{e_{\text{min}}}$,
  positive and negative \emph{infinity}, \emph{negative zero},
  and the \emph{not-a-number} error value.
\casio is mainly concerned with \emph{precision},
  not in avoiding exceptional values,
  so we will mainly not be concerned with these special values.
We write $\R(x)$ for the real number represented
  by the floating point number $x$.

The mantissa and exponent of a floating point value
  are fixed-width field; the width of these fields
  is set by the \emph{floating point number system} in use.
The floating point number system is a triple $(N_e, N_m, \F)$,
  where $N_e$ is the number of bits available for the exponent,
  and $N_m$ is the number of bits available for the mantissa.
$\F$ is a \emph{rounding mode}---%
  a function from real numbers to floating-point values
  such that $\F(\R(x)) = x$.
Rounding modes might round values
  not exactly representable as a floating-point value
  up, or may round to the closest floating-point value
  with a least-significant bit of zero,
  or any of a number of other rounding modes.
For example, IEEE double-precision allocates
  52 bits for the mantissa ($N_m = 52$)
  and 11 bits for the exponent ($N_e = 11$);
  together with a sign bit, this accounts for 64 bits.
It allows applications to choose between rounding up, down, or to zero;
  or rounding to the mathematically closer value, with ties breaking
  either toward the value with a zero least significant bit,
  or away from zero.
We denote by $\mathbf{FP}(N_e, N_m, \F)$
  the floating point number system with $N_e$ and $N_m$ as given.

Primitive arithmetic operators on floating point numbers
  such as addition and multiplication
  are guaranteed to produce accurate results.
For example, the floating point sum
  $x \oplus y$ of floating point values $x$ and $y$
  is guaranteed to be equal to the real-number sum
  of $x$ and $y$, rounded: $\F(\R(x) + \R(y))$.
Operators such as exponentiation, and trigonometric functions
  are implemented by libraries, and do not carry such guarantees.
An implementation of a mathematical function $f(x_1, x_2, \dotsc)$
  usually guarantees that its result
  is among the $u$ closest floating point values
  to the exact result $\F(f(\R(x_1), \R(x_2), \dotsc)))$.
$u$ is typically less than 8; that is,
  a library implementation usually guarantees that all
  but the two or three least-significant bits are correct.

\end{document}
